---
title: "Project Proposal"
output:
     html_document
---

__Proposal Deadline: Jan 30, 2017 11:59pm__


- Edit this markdown file by __replacing only__ what is there in the
chunks named `q0`, `q1`, `q2`, ..., `q5` with your team's appropriate
answers. These chunks have been filled in as an example.

- Place the filled in markdown in any one of the team member's private
directory with the same name `proposal.Rmd`. Do not change the name!


0. Who are the authors of this package? See format below.

```{r q0}
authors <- list(sunet1 = "John Peach",
                sunet2 = "Marilson Campos",
                sunet3 = "Prasad Telukuntla")
```

1. What is your package trying to achieve?  Who is it aimed at?

```q1

1. package name is testdata, This package is a tool/framework that validates input data just like how we validate the code. In some cases, Data scientists spend quiet a bit of time to retrain model when the input dataset changes. At that point, Data scientist has to inspect the changes in the data and see what caused model to fail or not work as expected such as distribution of features, outliers and missing values etc. This Framework helps define characteristics of dataset and validate them. Aditionally, this framework can plug right into preprocess stage of prediction pipeline.  

2. testdata package is aimed at Data scientists, Analytics and Data Engineers in any industry where the statistical/machine learning models are used/deployed in production (e.g., serving Data products, Apps etc to the customer etc).



```

2. Briefly describe the background if that's not obvious.  What field
   or area of interest is it related to?  (Just briefly, in a paragraph
   or two, maximum.) References, including web references, are welcome
   where available.

```q2

This package is a framework and applicable to most of the domains. Following is the scope of testdata validation framework.

1) Validate list of dataframes
2) Validate the Model objects (lm, GLM etc)
3) Validate list of row names
4) Validate list of all column names
5) Validate All factors present for a set of columns
6) Valdiate for a given set of values present in specified columns
7) Valdiate if all values are present for a set of columns
8) Validate if a particular column is Ascending order or Descending order
9) Metric (Mean, sum, standard deviation etc) are within a range
10) Custom Metric (with extra parameters for computing it) within a range
11) Expected Number of factors
12) All factors exit
13) Detecting Outliers based on the definition
14) Checking for Greater than, Greater than or equal, Less than and Less than or equal
15) Columns Floor and Ceiling
16) Test for normalcy
17) Vignette documenation of each of the use case


This package is inspired by the Hadley Wickham's package testthat. 
github: https://github.com/hadley/testthat
Reference: http://r-pkgs.had.co.nz/tests.html

```

3. Provide one or two use cases; that is, in a sentence or so say how
   you see someone using the package.

```q3

Data Scientists once they complete the model, they can run this pacakge on the new dataset that comes in to check if the data validation is passed before performing final prediction. This can also work as Monitorting component in the production pipeline. As data validation is hard and only Data Scientist can help define that through this pacakage.



```

4. If you can, list the main functions or other software you plan on
   including.  No need to implement anything now, but you should have
   an idea.

```q4

The R package will provide the following functions: reading and
writing data, creating graphs and visualizations, running
parameterized queries against datasets.  The R package will also
provide functions for generating vignette and report on the test results.

```

```{r, echo = FALSE, hide = TRUE}
none <- "none"
basic <- "basic"
proficient <- "proficient"
```

5. For languages C/C++, Java, Python, Ruby, please specify your
   proficiency: Proficient, basic knowledge, no knowledge.  (This
   helps us evaluate the proposal sometimes and offer suggestions.)

```{r q5}
languages <- list(
    ## replace the values below appropriately for each member
    sunet1 = c(cxx = proficient,
               java = proficient,
               python = proficient,
               ruby = none)
    ,
    sunet2 = c(cxx = proficient,
               java = proficient,
               python = proficient,
               ruby = none)
    ,
    sunet3 = c(cxx = proficient,
               java = proficient,
               python = proficient,
               ruby = none)
)
```








